---
title: "Common Data Set 2020-2021"
author: "California State University, East Bay"
date: "generated on `r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document: default
  word_document: default
  pdf_document: default
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhf{}
- \chead{Common Data Set 2020-2021}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r include=FALSE}
# Install and load packages

if(!require(tidyverse)) install.packages("tidyverse",repos = "http://cran.us.r-project.org")

library(tidyverse)
library(knitr)

rm(list = ls())

# Create a reference for the directory where the ERS submission files are stored.

ERSdir <- file.path("K:\\My Drive\\AAprojects\\CDSusingR\\2020-2021\\ERS\\")

# Two digit campus code.  East Bay's is 05.

cc <- "05"

# Save the following two files for CIP codes in Section J, 
#    commonc-cosarcipcodefxvw-prd-1.csv and
#    CIPCategory.csv
# and create a reference for the directory.

CIPdir <- file.path("K:\\My Drive\\AAprojects\\CDSusingR")

```


```{r include=FALSE}

# Read in fall ERS files for the enrollment sections, B1 and B2. The data are also used to find 2019 freshmen retention in section B22 and freshman profile, C9-C12. 
# I added self support students following the IPEDS guideline.

ERSS <- read.fwf(file.path(ERSdir, paste0("ERSS\\ERS.PROD.T20204",cc)), 
                 width = c(-24,1,-1,1,-13,6,-7,1,
                           1,1,-10,1,-39,2,
                           -3,3,3,3,3,
                           -3,2,2,-4,2,-76,1,
                           3,-13,1,-14,9,-22,3,
                           3,4,-24,2),
                 col.names=c('SexCode','CitizenshipCode','InstitutionOfOrigin','EnrollmentStatus',
                             'StudentLevelCode','DegreeObjectiveCode','StudentStandingCode','ACTWritingScore',
                             'LDunits','UDunits','GradUnits','HighSchoolGPA',
                             'ACTEnglishScore','ACTMathScore','ACTCompositeScore','MatriculationType',
                             'PreUnits','IPEDSEthnicity','EmployeeID', 'SATReadWriteScore',
                             'SATMathScore','SATCompositeScore','GenderIdentityCode'),
                 colClasses=c('character','character','character','numeric',
                              'numeric','numeric','character','numeric',
                              'numeric','numeric','numeric','numeric',
                              'numeric','numeric','numeric','character',
                              'numeric','numeric','character','numeric',
                              'numeric','numeric','numeric'))

sERS <- read.fwf(file.path(ERSdir, paste0("sERSS\\SER.PROD.T20204",cc)), 
                 width = c(-24,1,-1,1,-13,6,-7,1,
                           1,1,-10,1,-39,2,
                           -3,3,3,3,3,
                           -3,2,2,-4,2,-76,1,
                           -16,1,-5,9,-17,3,-9,3,
                           3,4,-24,2),
                 col.names=c('SexCode','CitizenshipCode','InstitutionOfOrigin','EnrollmentStatus',
                             'StudentLevelCode','DegreeObjectiveCode','StudentStandingCode','ACTWritingScore',
                             'LDunits','UDunits','GradUnits','HighSchoolGPA',
                             'ACTEnglishScore','ACTMathScore','ACTCompositeScore','MatriculationType',
                             'IPEDSEthnicity','EmployeeID','PreUnits', 'SATReadWriteScore',
                             'SATMathScore','SATCompositeScore','GenderIdentityCode'),
                 colClasses=c('character','character','character','numeric',
                              'numeric','numeric','character','numeric',
                              'numeric','numeric','numeric','numeric',
                              'numeric','numeric','numeric','character',
                              'numeric','character','numeric','numeric',
                              'numeric','numeric','numeric'))
ERS_Enr <- dplyr::bind_rows("R" = ERSS, "E" = sERS, .id = "FileType")
rm(ERSS,sERS)


# Read in degree data for total degrees conferred, B3, and by CIP codes, J.
# East Bay combines fall degree data into spring file. If you have a separate file for fall, you'll need to add it here.

ERSDsummer <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20203",cc)), 
                 width = c(-53,1,5),
                 col.names=c('DegreeLevelCode','MajorCode'),
                 colClasses=c('numeric','character'))
ERSDfall_spring <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20212",cc)), 
                     width = c(-53,1,5),
                     col.names=c('DegreeLevelCode','MajorCode'),
                     colClasses=c('numeric','character'))
ERSD <- rbind(ERSDsummer,ERSDfall_spring)
rm(ERSDsummer, ERSDfall_spring)


# Read 2014 enrollment data to define cohort for graduation rates in section B4.
# Chancellor's Office did not include self-support students for GI2025 2014 cohort, so I use the state side only.

ERSS_Coh <- read.fwf(file.path(ERSdir, paste0("ERSS\\ERS.PROD.T20144",cc)), 
                 width = c(-53,1,1,1,-10,1,-44,3,3,3,-93,3,-28,9),
                 col.names=c('EnrollmentStatus','StudentLevelCode','DegreeObjectiveCode','StudentStandingCode',
                             'LDunits','UDunits','GradUnits','PreUnits','EmployeeID'),
                 colClasses=c('numeric','numeric','numeric','character',
                              'numeric','numeric','numeric','numeric','character'))
ERSS_Coh <- mutate(ERSS_Coh, TotalUnits = (PreUnits+LDunits+UDunits+GradUnits)/10)


# Select first time full time undergraduate students from the cohort data.

FTFT_Coh <- subset(ERSS_Coh,EnrollmentStatus==5 & StudentLevelCode < 5 & TotalUnits >= 12, select = EmployeeID)
rm(ERSS_Coh)


# For 4,5,6 year graduation records, read in fall 2014 through summer 2020 data.
# Again, make sure to add fall degree data if your campus files it separately.

ERSD2152 <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20152",cc)), 
                 width = c(-9,5,-39,1,-50,9),
                 col.names=c('YearTerm','DegreeLevelCode','EmployeeID'),
                 colClasses=c('numeric','character','character'))
ERSD2153 <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20153",cc)), 
                 width = c(-9,5,-39,1,-50,9),
                 col.names=c('YearTerm','DegreeLevelCode','EmployeeID'),
                 colClasses=c('numeric','character','character'))
ERSD2162 <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20162",cc)), 
                 width = c(-9,5,-39,1,-50,9),
                 col.names=c('YearTerm','DegreeLevelCode','EmployeeID'),
                 colClasses=c('numeric','character','character'))
ERSD2163 <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20163",cc)), 
                 width = c(-9,5,-39,1,-50,9),
                 col.names=c('YearTerm','DegreeLevelCode','EmployeeID'),
                 colClasses=c('numeric','character','character'))
ERSD2172 <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20172",cc)), 
                 width = c(-9,5,-39,1,-50,9),
                 col.names=c('YearTerm','DegreeLevelCode','EmployeeID'),
                 colClasses=c('numeric','character','character'))
ERSD2173 <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20173",cc)), 
                 width = c(-9,5,-39,1,-50,9),
                 col.names=c('YearTerm','DegreeLevelCode','EmployeeID'),
                 colClasses=c('numeric','character','character'))
ERSD2182 <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20182",cc)), 
                 width = c(-9,5,-39,1,-50,9),
                 col.names=c('YearTerm','DegreeLevelCode','EmployeeID'),
                 colClasses=c('numeric','character','character'))
ERSD2183 <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20183",cc)), 
                 width = c(-9,5,-39,1,-50,9),
                 col.names=c('YearTerm','DegreeLevelCode','EmployeeID'),
                 colClasses=c('numeric','character','character'))
ERSD2192 <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20192",cc)), 
                 width = c(-9,5,-39,1,-50,9),
                 col.names=c('YearTerm','DegreeLevelCode','EmployeeID'),
                 colClasses=c('numeric','character','character'))
ERSD2193 <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20193",cc)), 
                 width = c(-9,5,-39,1,-50,9),
                 col.names=c('YearTerm','DegreeLevelCode','EmployeeID'),
                 colClasses=c('numeric','character','character'))
ERSD2202 <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20202",cc)), 
                 width = c(-9,5,-39,1,-50,9),
                 col.names=c('YearTerm','DegreeLevelCode','EmployeeID'),
                 colClasses=c('numeric','character','character'))
ERSD2203 <- read.fwf(file.path(ERSdir, paste0("ERSD\\ERD.PROD.T20203",cc)), 
                 width = c(-9,5,-39,1,-50,9),
                 col.names=c('YearTerm','DegreeLevelCode','EmployeeID'),
                 colClasses=c('numeric','character','character'))
ERSDtotal <- rbind(ERSD2152,ERSD2153,ERSD2162,ERSD2163,ERSD2172,ERSD2173,ERSD2182,ERSD2183,ERSD2192,ERSD2193,ERSD2202,ERSD2203) 
rm(ERSD2152,ERSD2153,ERSD2162,ERSD2163,ERSD2172,ERSD2173,ERSD2182,ERSD2183,ERSD2192,ERSD2193,ERSD2202,ERSD2203)


# Read in fall 2019 enrollment files for the section B22, one year retention.

ERSS_ret <- read.fwf(file.path(ERSdir, paste0("ERSS\\ERS.PROD.T20194",cc)), 
                 width = c(-24,1,-1,1,-26,1,1,1,-10,1,-44,3,3,3,-92,1,3,-13,1,-14,9,-56,2),
                 col.names=c('SexCode','CitizenshipCode','EnrollmentStatus','StudentLevelCode',
                             'DegreeObjectiveCode','StudentStandingCode','LDunits','UDunits','GradUnits',
                             'MatriculationType','PreUnits','IPEDSEthnicity','EmployeeID','GenderIdentityCode'),
                 colClasses=c('character','character','numeric','numeric',
                              'numeric','character','numeric','numeric','numeric',
                              'character','numeric','numeric','character','numeric'))
sERS_ret <- read.fwf(file.path(ERSdir, paste0("sERSS\\SER.PROD.T20194",cc)), 
                 width = c(-24,1,-1,1,-26,1,1,1,-10,1,-44,3,3,3,-92,1,-16,1,-5,9,-13,3,-46,2),
                 col.names=c('SexCode','CitizenshipCode','EnrollmentStatus','StudentLevelCode',
                             'DegreeObjectiveCode','StudentStandingCode','LDunits','UDunits','GradUnits',
                             'MatriculationType','IPEDSEthnicity','EmployeeID','PreUnits','GenderIdentityCode'),
                 colClasses=c('character','character','numeric','numeric',
                              'numeric','character','numeric','numeric','numeric',
                              'character','numeric','character','numeric','numeric'))
ERS_ret <- dplyr::bind_rows("R" = ERSS_ret, "E" = sERS_ret, .id = "FileType")
rm(ERSS_ret, sERS_ret)


# Read in application data for C1 and D2.

ERSAsummer <- read.fwf(file.path(ERSdir, paste0("ERSA\\ERA.PROD.T20203",cc)), 
                 width = c(-27,1,-43,1,1,-20,1
                           ,1,-239,9,-54,2),
                 col.names=c('AdmissionStatus','DegreeObjectiveCode','EnrollmentStatus','SexCode',
                             'StudentLevelCode','EmployeeID','GenderIdentityCode'),
                 colClasses=c('character','numeric','numeric','character',
                              'numeric','character','numeric'))
ERSAfall <- read.fwf(file.path(ERSdir, paste0("ERSA\\ERA.PROD.T20204",cc)), 
                 width = c(-27,1,-43,1,1,-20,1
                           ,1,-239,9,-54,2),
                 col.names=c('AdmissionStatus','DegreeObjectiveCode','EnrollmentStatus','SexCode',
                             'StudentLevelCode','EmployeeID','GenderIdentityCode'),
                 colClasses=c('character','numeric','numeric','character',
                              'numeric','character','numeric'))

ERSA <- rbind(ERSAsummer, ERSAfall)
rm(ERSAsummer,ERSAfall)


```


# A. GENERAL INFORMATION

### A0. Respondent Information (Not for Publication)

Respondent      | Information
----------------|----------------------------------------------       
Name:           | Your name here    
Title:          | Research Associate    
Office:         | Institutional Effectiveness and Research  
Mailing Address:| 25800 Carlos Bee Blvd    
City/State/Zip: | Hayward, CA 94542  
Phone:          | 510-885-3000   
Fax:            |    
Email Address:  |    


### A1. Address Infomration

Address                            | Information
-----------------------------------|--------------------------------------------
Name of University:                | California State University, East Bay 
Mailing Address:                   | 25800 Carlos Bee Blvd. 
City/State/Zip:                    | Hayward, CA 94542 
Main Phone Number:                 | 714-278-2011 
WWW Home Page Address:             | http://www.csueastbay.edu 
Admissions Phone Number:           |  
Admissions Toll-Free Number:       | 
Admissions Office Mailing Address: |  	
City/State/Zip:                    | Hayward, CA 94542 
Admissions Fax Number:             | 
Admissions E-mail Address:         | admissions\@csueastbay.edu 

### A2. Source of Institutional Control
- [x] Public
- [ ] Private (nonprofit)
- [ ] Proprietary
    
### A3. Classify your undergraduate institution
- [x] Coeducational college
- [ ] Men's college
- [ ] Women's college
    
### A4. Academic year calendar

   Academic year         | calendar
-------------------------|--------------------------------------
- [x] Semester           | - [ ] 4-1-4
- [ ] Quarter            | - [ ] Continuous
- [ ] Trimester          | - [ ] Differs by program (describe:)
- [ ] Other (describe:)  |

### A5. Degrees offered by your institution

Degrees           | Offered
------------------|--------------------------------------
- [ ] Certificate | - [ ] Postbachelor's certificate 
- [ ] Diploma     | - [x] Master's 
- [ ] Associate   | - [ ] Post-master's certificate 
- [ ] Transfer    | - [ ] Doctoral degree-research 
- [ ] Terminal    | - [ ] Doctoral degree-professional 
- [x] Bachelor's  | - [X] Doctoral degree-other 





# B. ENROLLMENT AND PERSISTENCE

### B1.	Institutional Enrollment - Men and Women

```{r echo=FALSE}


# Using the fall enrollment data, Add all units to determine full time part time status.

ERS_Enr <- mutate(ERS_Enr, Units = (PreUnits+LDunits+UDunits+GradUnits)/10)
ERS_Enr <- 
  ERS_Enr %>%
    group_by(EmployeeID) %>% 
    mutate(TotalUnits=sum(Units))

# Check for students enrolled as state and self.

DupeTable <- data.frame(table(ERS_Enr$EmployeeID))
#DupeTable[DupeTable$Freq > 1,]  

Check <- ERS_Enr[ERS_Enr$EmployeeID %in% DupeTable$Var1[DupeTable$Freq > 1],14:17]
#Check[order(Check$EmployeeID),]

#Select records based on Matriculation Type. 

colnames(DupeTable) <- c("EmployeeID","N")
ERS_Enr <- merge(ERS_Enr, DupeTable, by="EmployeeID")

Singles <- subset(ERS_Enr, N==1)
Doubles <- subset(ERS_Enr,(N==2 & MatriculationType==FileType))
rm(ERS_Enr)

ERS_Enr <- rbind(Singles,Doubles)
rm(DupeTable, Check, Singles, Doubles)

## N in ERS Sex Code needs to be redistributed to either W or M using Gender Identity Code or last digit of ID number.

LastDigID <- strtoi(substr(ERS_Enr$EmployeeID,9,9))
ERS_Enr <- cbind(ERS_Enr,LastDigID)
Gender <- recode(ERS_Enr$SexCode, "F"="W", "M"="M")
ERS_Enr <- cbind(ERS_Enr,Gender)
rm(LastDigID,Gender)
CDSsex <- with(ERS_Enr, ifelse(SexCode=="N", 
                            ifelse(GenderIdentityCode>=14,ifelse(LastDigID%%2==0,"W","M"),
                                   ifelse(GenderIdentityCode %in% c(10,12),"M","W")) ,Gender))
ERS_Enr <- cbind(ERS_Enr,CDSsex)

## Recode ethnicity to match the order, then isolate non residents using Citizenship Code.

Ethnicity <- recode(ERS_Enr$IPEDSEthnicity,"7"=2,"2"=3,"1"=4,"3"=5,"4"=6,"5"=7,"6"=8,"8"=9)
ERS_Enr <-cbind(ERS_Enr,Ethnicity)
rm(Ethnicity)
CDSEthnicity <- with(ERS_Enr, ifelse(CitizenshipCode %in% c("F","J","O","N"),1,Ethnicity ))

## Here degree type is defined using Enrollment Status and Student Level Code, and Career based on Level and Degree Objective Code.

CDSdegreeType <- with(ERS_Enr, ifelse(EnrollmentStatus == 5, ifelse(StudentLevelCode < 5,1,5), 
                                   ifelse(EnrollmentStatus==6, ifelse(StudentLevelCode<5,4,7),
                                          ifelse(StudentLevelCode==1,2,ifelse(StudentLevelCode==5,6,3)) )))
CDScareer <- with(ERS_Enr, ifelse(StudentLevelCode < 5 | (DegreeObjectiveCode < 5 & DegreeObjectiveCode > 0), 1, 2))
ERS_Enr <- cbind(ERS_Enr,CDScareer)
rm(CDScareer)

FTPT <- with(ERS_Enr, ifelse((CDScareer == 1 & TotalUnits >= 12)|(CDScareer ==2 & TotalUnits >= 9), 'FT', 'PT'))
B1 <- ftable(CDSdegreeType, FTPT, CDSsex, row.vars = "CDSdegreeType")
rm(FTPT, CDSsex)

sec1 <- B1[1:3,1:4]
sec1wt <- rbind(sec1,c(colSums(sec1)))
sec2 <- rbind(c(colSums(sec1)),B1[4,1:4])
sec2wt <- rbind(sec2,c(colSums(sec2)))
sec3 <- B1[5:7,1:4]
sec3wt <- rbind(sec3,c(colSums(sec3,na.rm = TRUE)))
sec4 <- colSums(B1)
rm(B1)
B1 <- rbind(sec1,sec2wt,sec3wt,sec4)
rm(sec1, sec1wt, sec2, sec2wt, sec3, sec3wt, sec4)

options(scipen = 999)
totalUG <- sum(B1[6,1:4])
totalG <- sum(B1[10,1:4])
grandTotal <- totalUG + totalG
rownames(B1) <- c("Degree-seeking, first-time freshman","Degree-seeking, Other first-year","Degree-seeking, Other",
                     "Total degree-seeking","All other undergraduates","Total Undergraduates",
                     "Degree-seeking, first-time grad","Degree-seeking, Other grad","All other grad","Total grad","Total all students")
colnames(B1) <- c("FT Men","FT Women","PT Men","PT Women")
kable(B1)
```

Total all undergraduates: `r totalUG`   
Total all graduate: `r totalG`   
GRAND TOTAL ALL STUDENTS: `r grandTotal`    

### B2.	Enrollment by Racial/Ethnic Category

```{r echo=FALSE}

# Clean up the data no longer needed.
rm(totalUG, totalG, grandTotal)

# Use the fields defined above.

B2table <- table(CDSEthnicity, CDSdegreeType) 
rm(CDSEthnicity, CDSdegreeType)

B2sec2 <- rowSums(B2table[,1:3])
B2sec3 <- rowSums(B2table[,1:4])
B2cols <- cbind(B2table[1:9,1],B2sec2,B2sec3)
B2total <- colSums(B2cols)
B2 <- rbind(B2cols,B2total)
rm(B2sec2, B2sec3, B2cols, B2total, B2table)

rownames(B2) <- c("Nonresident aliens","Hispanic/Latino","Black or African American, non-Hispanic","White, non-Hispanic",
                  "American Indian or Alaska Native, non-Hispanic","Asian, non-Hispanic","Native Hawaiian or other Pacific Islander, non-Hispanic",
                  "Two or more races, non-Hispanic","Race and/or ethnicity unknown","TOTAL")
colnames(B2) <- c("Degree-Seeking First-Time First Year","Degree-Seeking Undergraduates","Total Undergraduates")

kable(B2)

```

## Persistence
### B3.	Number of degrees awarded by your institution from July 1, 2019, to June 30, 2020

```{r echo=FALSE}

rm(B2)

# Using degree data, group degree levels.

Degree <- recode(ERSD$DegreeLevelCode,"2"="BA","3"="BA","4"="BA","5"="MA","6"="MA","7"="MA","8"="Doc")
B3table <- table(Degree)
rm(Degree)

ba <- B3table[1]
ma <- B3table[3]
dr <- B3table[2]
rm(B3table)

#### NOTE1: CSUEB does not offer certificates or associate degrees.   
#### NOTE2: CSUEB offers only one doctoral degree (EDD) and it is considered to be other doctoral degree.  

```
   
Certificate/diploma: 0     
Associate degrees: 0     
Bachelor's degrees: `r ba`     
Postbachelor's certificates: 0     
Master's degrees: `r ma`     
Post-master's certificates: 0     
Doctoral degrees - research/scholarship: 0     
Doctoral degrees - professional practice: 0     
Doctoral degrees - other:  `r dr`  


## B4-B21 Graduation Rates
### For Bachlor's Programs using Fall 2014 Cohort

```{r echo=FALSE}

rm(ba, ma, dr)

# I will omit the breakdown by Pell grant and Stafford loans, as they are not available from ERS files. 

# select only UG degrees in case a higher degree is earned during the same time frame.
UGdegrees <- subset(ERSDtotal,DegreeLevelCode < 5)

# match cohort with UG degree records.
Graduated <- merge(FTFT_Coh,UGdegrees,by="EmployeeID")
rm(UGdegrees, ERSDtotal)

# check for multiple degrees
#Graduated[duplicated(Graduated$EmployeeID)]

FourOrless <- with(Graduated, ifelse(YearTerm<=20183,1,0))
FourToFive <- with(Graduated, ifelse(YearTerm<=20193,ifelse(YearTerm<=20183,0,1),0))
FiveToSix <- with(Graduated, ifelse(YearTerm<=20203,ifelse(YearTerm<=20193,0,1),0))
SixOrLess <- with(Graduated, ifelse(YearTerm<=20203,1,0))
rm(Graduated)

CohortN <- nrow(FTFT_Coh)
FourOrLessN <- sum(FourOrless)
FourToFiveN <- sum(FourToFive)
FiveToSixN <- sum(FiveToSix)
SixOrLessN <- sum(SixOrLess)
SixYrGR <- round((SixOrLessN/CohortN)*100,digits = 0)
rm(FourOrless, FourToFive, FiveToSix, SixOrLess, FTFT_Coh)

```

A. Initial 2014 Cohort of first-time full-time bachelor's degree-seeking undergraduate students: `r CohortN`     
B. Of the initial 2014 cohort, how many did not persist: 0   
C. Final 2014 cohort, after adjusting for allowable exclusions: `r CohortN`   
D. Completed in four years or less: `r FourOrLessN`   
E. Completed in more than four years but in five years or less: `r FourToFiveN`   
F. Completed in more than five years but in six years or less: `r FiveToSixN`   
G. Total graduating within six years: `r SixOrLessN`   
H. Six year graduation rate: `r SixYrGR`   

## B22 Retention Rates
### One Year Retention Rate of Fall 2019 full-time bachelor's degree-seeking undergraduate students

```{r echo=FALSE}

rm(CohortN, FourOrLessN, FourToFiveN, FiveToSixN, SixOrLessN, SixYrGR)

# add all the units together.
ERS_ret <- mutate(ERS_ret, Units = (PreUnits+LDunits+UDunits+GradUnits)/10)
ERS_ret <- 
  ERS_ret %>%
    group_by(EmployeeID) %>% 
    mutate(TotalUnits=sum(Units))

# Check for students enrolled as state and self.

DupeTable <- data.frame(table(ERS_ret$EmployeeID))
#DupeTable[DupeTable$Freq > 1,]  

Check <- ERS_ret[ERS_ret$EmployeeID %in% DupeTable$Var1[DupeTable$Freq > 1],12:15]
#Check[order(Check$EmployeeID),]

#Select records based on Matriculation Type to unduplicate. 

colnames(DupeTable) <- c("EmployeeID","N")
ERS_ret <- merge(ERS_ret, DupeTable, by="EmployeeID")

Singles <- subset(ERS_ret, N==1)
Doubles <- subset(ERS_ret,(N==2 & MatriculationType==FileType))
rm(ERS_ret)

ERS_ret <- rbind(Singles,Doubles)
rm(DupeTable, Check, Singles, Doubles)

FTFT_ret <- subset(ERS_ret,EnrollmentStatus==5 & StudentLevelCode < 5 & TotalUnits >= 12, select= EmployeeID)

# match Cohort19 with fall 20 enrollment records.

RetainedList <- merge(FTFT_ret,ERS_Enr,by="EmployeeID")
Retained <- sprintf("%1.0f%%" ,(nrow(RetainedList)/nrow(FTFT_ret))*100)
rm(RetainedList, FTFT_ret)

```

B22: Percent of fall 2019 cohort enrolled in fall 2020:`r Retained`


# C. FIRST-TIME, FIRST-YEAR (FRESHMAN) ADMISSION

## C1-C2: Applications

### C1.	First-time, first-year (freshman) students: 

```{r echo=FALSE}

rm(Retained)


# Using application data, define sex.

LastDigID <- strtoi(substr(ERSA$EmployeeID,9,9))
ERSA <- cbind(ERSA,LastDigID)
Gender <- recode(ERSA$SexCode, "F"="W", "M"="M")
ERSA <- cbind(ERSA,Gender)
rm(LastDigID,Gender)
CDSsex <- with(ERSA, ifelse(SexCode=="N", 
                            ifelse(GenderIdentityCode>=14,ifelse(LastDigID%%2==0,"W","M"),
                                   ifelse(GenderIdentityCode %in% c(10,12),"M","W")) ,Gender))
ERSA <- cbind(ERSA,CDSsex)
rm(CDSsex)

# Group Admission Status.

ERSA$AdmissionStatus <- as.factor(ERSA$AdmissionStatus)

levels(ERSA$AdmissionStatus)<-list(
                   'Admitted' = c('A','C','F','H','N','P','G'), 
                   'Not Admitted' = c('D','T','U','J','K') ,
                   'Incomplete' = c('I','M'),
                   'No Decision' = c('E','W','Q'))

# select new undergraduate students and exclude incomplete applications.

ERSA <- subset(ERSA, (EnrollmentStatus == 4 | EnrollmentStatus == 5) & StudentLevelCode < 5 & AdmissionStatus != 'Incomplete')

# Define First time freshman (or transfer for Section D2).

# Check they are all degree seeking.
#table(ERSA$DegreeObjectiveCode)

Freshman_App <- subset(ERSA, EnrollmentStatus == 5)
Transfer_App <- subset(ERSA, EnrollmentStatus == 4)


appliedM <- table(Freshman_App$CDSsex)[1]
appliedW <- table(Freshman_App$CDSsex)[2]
admitM <- table(Freshman_App$AdmissionStatus,Freshman_App$CDSsex)[1,1]
admitW <- table(Freshman_App$AdmissionStatus,Freshman_App$CDSsex)[1,2]

# First time freshman fall enrollment (from Section B).

enrollFTM <- B1[1,1]
enrollPTM <- B1[1,2]
enrollFTW <- B1[1,3]
enrollPTW <- B1[1,4]

```

Total first-time, first-year (freshman) men who applied: `r appliedM` .    
Total first-time, first-year (freshman) women who applied: `r appliedW` .    

Total first-time, first-year (freshman) men who were admitted: `r admitM` .    
Total first-time, first-year (freshman) women who were admitted: `r admitW` .    

Total full-time, first-time, first-year (freshman) men who enrolled: `r enrollFTM` .    
Total part-time, first-time, first-year (freshman) men who enrolled: `r enrollPTM` .    

Total full-time, first-time, first-year (freshman) women who enrolled: `r enrollFTW` .    
Total part-time, first-time, first-year (freshman) women who enrolled: `r enrollPTW` .  

### C2.	Freshman wait-listed students

Do you have a policy of placing students on a waiting list?
- [ ] Yes
- [x] No

Is your waiting list ranked?
- [ ] Yes
- [ ] No

## C3-C5: Admission Requirements
### C3. High school completion requirement

- [x] High school diploma is required and GED is accepted
- [ ] High school diploma is required and GED is not accepted
- [ ] High school diploma or equivalent is not required  
    
### C4. Does your institution require or recommend a general college-preparatory program for degree-seeking students?

- [x] Require
- [ ] Recommend
- [ ] Neither require nor recommend   
		
### C5. Distribution of high school units required and/or recommended.

Subject                          |Units Required | Units Recommended   
---------------------------------|---------------|--------------------
Total academic units             |      15       |       16   
English                          |       4       |        4   
Mathematics                      |       3       |        3   
Science                          |       2       |        2   
of these, units that must be lab |       2       |        2   
Foreign language                 |       2       |        3   
Social studies                   |       1       |        1   
History                          |       1       |        1   
Academic electives               |       1       |        1   
Computer Science                 |       0       |        0   
Visual/Performing Arts           |       1       |        1   
Other (specify)                  |       0       |        0   

## C6-C7: Basis for Selection
### C6. Open Admission Policy

- [ ] Yes
- [X] No
    
### C7. Admission decisions.

Subject                          |Very Important|Important|Considered|Not Considered   
---------------------------------|--------------|---------|----------|----------------
Academic                         |
Rigor of secondary school record |              |         |          |       x   
Class rank                       |              |         |          |       x   
Academic GPA                     |      x       |         |          |   
Standardized test scores         |      x       |         |          |   
Application Essay                |              |         |          |       x   
Recommendation(s)                |              |         |          |       x   
---------------------------------|--------------|---------|----------|-----------------
Nonacademic                      |
Interview                        |              |         |          |       x   
Extracurricular activities       |              |         |          |       x   
Talent/ability                   |              |         |          |       x   
Character/personal qualities     |              |         |          |       x   
First generation                 |              |         |          |       x   
Alumni/ae relation               |              |         |          |       x   
Geographical residence           |      x       |         |          |   
State residency                  |      x       |         |          |   
Religious affiliation/commitment |              |         |          |       x   
Racial/ethnic status             |              |         |          |       x   
Volunteer work                   |              |         |          |       x   
Work experience                  |              |         |          |       x   
Level of applicant's interest    |              |         |          |       x   


## C8: SAT and ACT policies

### Entrance exams
Does your institution make use of SAT, ACT, or SAT Subject Test scores in admission decisions for first-time, first-year, degree-seeking applicants?   

- [x] Yes
- [ ] No

#### C8A. SAT/ACT requirement
Subject                          |Require|Recommend|Consider if Submitted|Not Used   
---------------------------------|-------|---------|---------------------|---------
SAT or ACT                       |   x   |         |                     |   
ACT only                         |       |         |          x          |   
SAT only                         |   x   |         |                     |   
SAT and SAT Subject Tests or ACT |       |         |          x          |   
SAT Subject Tests only           |       |         |          x          |   
  		
#### C8B. SAT/ACT writing requirement
- [ ] Yes
- [x] No

#### C8C. SAT/ACT essary requirement 
Essay required                             	 | SAT essay | ACT essay   
---------------------------------------------|-----------|-----------
For admission                                |           |   
For placement                                |           |   
For advising                                 |           |   
In place of an application essay             |           |   
As a validity check on the application essay |           |   
No college policy as of now                  |    x      |     x   
Not using essay component                    |           |   

#### C8D. SAT/ACT scores for advising
- [ ] Yes
- [x] No

#### C8E. SAT/ACT deadline for fall admission
January 15 

#### C8F. Clarification of test policies

#### C8G. Please indicate which tests your institution uses for placement

- [ ] SAT
- [ ] ACT
- [ ] SAT Subject Test
- [ ] AP
- [ ] CLEP
- [ ] Institutional Exam
- [ ] State Exam (Specify):


## C9-C12: Freshman Profile

### C9. SAT/ACT submission of Fall 2020 first-time, first-year (freshman) students

```{r echo=FALSE}

rm(B1, admitM, admitW, appliedM, appliedW, enrollFTM, enrollFTW, enrollPTM, enrollPTW)

# Number and percent scores submitted.

# Enrolled degree-seeking first time first year freshman.

Freshman_Enr <- subset(ERS_Enr, EnrollmentStatus == 5 & StudentLevelCode < 5)
Total <- nrow(Freshman_Enr)

# SAT and ACT submissions. I see ACT is 0 and SAT is blank when the scores are missing in our file. So skip na in SAT and 0 in ACT.

satN <- sum(!is.na(Freshman_Enr$SATCompositeScore))
actN <- colSums(Freshman_Enr != 0, na.rm = TRUE)[17]
SATpct <- sprintf("%1.0f%%",round((satN/Total)*100))
ACTpct <- sprintf("%1.0f%%",round((actN/Total)*100))
SAT <- c(SATpct,satN)
ACT <- c(ACTpct,actN)
C9a <- rbind(SAT,ACT)
colnames(C9a) <- c("Percent","Number")
rownames(C9a) <- c("Submitting SAT Scores","Submitting ACT Scores")
rm(Total, satN, actN, SATpct, ACTpct, SAT, ACT)

# SAT and ACT percentiles.

SATcomp <- quantile(Freshman_Enr$SATCompositeScore, c(.25, .75), na.rm = TRUE)
SATrw <- quantile(Freshman_Enr$SATReadWriteScore, c(.25, .75), na.rm = TRUE)
SATmath <- quantile(Freshman_Enr$SATMathScore, c(.25, .75), na.rm = TRUE)

ACTcomp <- quantile(Freshman_Enr$ACTCompositeScore[Freshman_Enr$ACTCompositeScore != 0], c(.25, .75))
ACTmath <- quantile(Freshman_Enr$ACTMathScore[Freshman_Enr$ACTMathScore != 0], c(.25, .75))
ACTeng <- quantile(Freshman_Enr$ACTEnglishScore[Freshman_Enr$ACTEnglishScore != 0], c(.25, .75))
ACTwrt <- quantile(Freshman_Enr$ACTWritingScore[Freshman_Enr$ACTWritingScore != 0], c(.25, .75))

C9b <- rbind(SATcomp, SATrw, SATmath, ACTcomp, ACTmath, ACTeng, ACTwrt)  
colnames(C9b) <- c("25th Percentile","75th Percentile")
rownames(C9b) <- c("SAT Composite","SAT Reading and Writing","SAT Math","ACT Composite","ACT Math","ACT English","ACT Writing")
rm(SATcomp, SATrw, SATmath, ACTcomp, ACTmath, ACTeng, ACTwrt)

# SAT Score ranges.

# SAT Reading and Writing and Math score tables are reported in one table.

# Read in SAT scores as frequency tables.
SATrw <- data.frame(table(cut(Freshman_Enr$SATReadWriteScore,
               breaks = c(200,299,399,499,599,699,800),
               labels = c("200-299","300-399","400-499","500-599","600-699","700-800"),
               include.lowest = TRUE)))

# Add percentages to the contingency table.
SATrw$Percent <- (SATrw$Freq/sum(SATrw$Freq, na.rm = TRUE))*100


SATmath <- data.frame(table(cut(Freshman_Enr$SATMathScore,
                     breaks = c(200,299,399,499,599,699,800),
                     labels = c("200-299","300-399","400-499","500-599","600-699","700-800"),
                     include.lowest = TRUE)))
SATmath$Percent <- (SATmath$Freq/sum(SATmath$Freq, na.rm = TRUE))*100

# Join reading&writing and math tables together.

SAT1 <- merge(SATrw, SATmath, by="Var1")
rm(SATrw, SATmath)

# Scores are repoted in descending order.

SAT1 <- SAT1[order(SAT1$Var1, decreasing = TRUE),]

# Add colum totals.

SAT1$Var1 <- as.character(SAT1$Var1)
SAT1 <- rbind(SAT1, c("Total", colSums(SAT1[,2:5])))

# Format percentages.

SAT1$Percent.x <- sprintf("%.1f%%", as.numeric(SAT1$Percent.x))
SAT1$Percent.y <- sprintf("%.1f%%", as.numeric(SAT1$Percent.y))

# Remove frequencies as only percentages are needed.

C9c <- SAT1[ -c(2,4) ]
rm(SAT1)

# Add column labels.

colnames(C9c) <- c("Score Range","SAT Reading and Writing","SAT Math")


# SAT Composite score table.

SATcomp <- data.frame(table(cut(Freshman_Enr$SATCompositeScore,
                     breaks = c(400,599,799,999,1199,1399,1600),
                     labels = c("400-599","600-799","800-999","1000-1199","1200-1399","1400-1600"),
                     include.lowest = TRUE)))
SATcomp$Percent <- (SATcomp$Freq/sum(SATcomp$Freq))*100

SAT2 <- SATcomp[order(SATcomp$Var1, decreasing = TRUE),]

SAT2$Var1 <- as.character(SAT2$Var1)
SAT2 <- rbind(SAT2, c("Total", colSums(SAT2[ ,c(2:3)])))
rm(SATcomp)

SAT2$Percent <- sprintf("%.1f%%", as.numeric(SAT2$Percent))

C9d <- SAT2[ -c(2) ]
rm(SAT2)
colnames(C9d) <- c("Score Range","SAT Composite")


# ACT score range table.

ACT_comp <- data.frame(table(cut(Freshman_Enr$ACTCompositeScore,
                      breaks = c(1,6,11,17,23,29,36),
                      labels = c("below 6","6-11","12-17","18-23","24-29","30-36"),
                      include.lowest = TRUE )))
ACT_comp$Percent <- (ACT_comp$Freq/sum(ACT_comp$Freq))*100

ACT_Eng <- data.frame(table(cut(Freshman_Enr$ACTEnglishScore,
                      breaks = c(1,6,11,17,23,29,36),
                      labels = c("below 6","6-11","12-17","18-23","24-29","30-36"),
                      include.lowest = TRUE )))
ACT_Eng$Percent <- (ACT_Eng$Freq/sum(ACT_Eng$Freq))*100

ACT_Math <- data.frame(table(cut(Freshman_Enr$ACTMathScore,
                      breaks = c(1,6,11,17,23,29,36),
                      labels = c("below 6","6-11","12-17","18-23","24-29","30-36"),
                      include.lowest = TRUE )))
ACT_Math$Percent <- (ACT_Math$Freq/sum(ACT_Math$Freq))*100

# Join the three tables together.

ACT_list <- list(ACT_comp, ACT_Eng, ACT_Math)
ACT <- Reduce(function(x, y) merge(x, y, by="Var1"), ACT_list, accumulate=FALSE)
rm(ACT_comp, ACT_Eng, ACT_Math, ACT_list)

# Order the ranges in descending order.

ACT <- ACT[order(ACT$Var1, decreasing = TRUE),]

# Add column totals.

ACT$Var1 <- as.character(ACT$Var1)
ACT <- rbind(ACT, c("Total", colSums(ACT[ , c(2:7)])))

# Format percentages.

ACT$Percent.x <- sprintf("%.1f%%", as.numeric(ACT$Percent.x))
ACT$Percent.y <- sprintf("%.1f%%", as.numeric(ACT$Percent.y))
ACT$Percent <- sprintf("%.1f%%", as.numeric(ACT$Percent))

# Only percentages.

C9e <- ACT[ -c(2,4,6) ]
rm(ACT)

colnames(C9e) <- c("Score Range","ACT Composite","ACT English","ACT Math")

kable(C9a)
kable(C9b)
kable(C9c, row.names = FALSE)
kable(C9d, row.names = FALSE)
kable(C9e, row.names = FALSE)
```

### C10. High school class rank 

### C11. High school GPA distribution.

```{r echo=FALSE}

rm(C9a, C9b, C9c, C9d, C9e)

# Score ranges are up to 5.0.  Not sure how to convert to 4.0 scale. 0 are excluded by making the lowest bound to be 1.

HS_GPA <- data.frame(table(cut(Freshman_Enr$HighSchoolGPA,
               breaks = c(1,100,199,249,299,324,349,374,399,Inf),
               labels = c("below 1.00","1.00-1.99","2.00-2.49","2.50-2.99","3.00-3.24","3.25-3.49","3.50-3.74","3.75-3.99"," 4.0+"),
               include.lowest = TRUE)))
HS_GPA$Percent <- (HS_GPA$Freq/sum(HS_GPA$Freq))*100

HS_GPA <- HS_GPA[order(HS_GPA$Var1, decreasing = TRUE),]

HS_GPA$Var1 <- as.character(HS_GPA$Var1)
HS_GPA <- rbind(HS_GPA, c("Total", colSums(HS_GPA[,2:3, drop=FALSE])))

HS_GPA$Percent <- sprintf("%.1f%%", as.numeric(HS_GPA$Percent))

C11 <- HS_GPA[-2]
rm(HS_GPA)
colnames(C11) <- c("Range","Percent")

kable(C11, row.names = FALSE)
```
   

### C12. Average high school GPA

```{r echo=FALSE}

rm(C11)

HSGPA_vector <- Freshman_Enr$HighSchoolGPA/100

# For the mean calculation, I converted 0 into missing.
HSGPA_vector[HSGPA_vector ==0] <- NA    
HSGPA_Mean <- round(mean(HSGPA_vector, na.rm = TRUE ), digits = 2)

HSGPA_Percent <- round((sum(!is.na(HSGPA_vector))/nrow(Freshman_Enr)*100),2)

rm(HSGPA_vector)

```
Average GPA of all degree-seeking, first-time, first-year (freshman) students: `r HSGPA_Mean`.

Percent of total First-time, first-year (freshman) students who submitted high school GPA: `r HSGPA_Percent`.

## C12-C13: Admission Policies

### C13. Application Fee

Does your institution have an application fee?  [x] Yes  [ ] No   
Amount of application fee: $55   
Can it be waived for applicants with financial need?  [x] Yes  [ ] No 

On-line application   
Same fee:  [x]   
Free:  [ ]   
Reduced:  [ ]   
Can it be waived for applicants with financial need? [x] Yes [ ] No   
	
### C14. Application Closing Date

Does your institution have an application closing date?  [x] Yes  [ ] No      
Application closing date (fall):  November 30   
Priority date:  October 30   

### C15. 

Are first-time, first-year students accepted for terms other than the fall? [ ] Yes  [x] No   


### C16. Notification of admission decision

On a rolling basis beginning (date):     
By (date):     
Other:      

### C17. Reply policy for admitted applicants

Must reply by (date): May 15     
Deadline for housing deposit (MMDD): 04/24    
Amount of housing deposit: $350    
Refundable if student does not enroll? Yes, in full    

### C18. Deferred admission

Does your institution allow students to postpone enrollment after admission? [ ] Yes [x] No  

### C19. Early admission

Does your institution allow high school students to enroll as full-time, firsttime, first-year (freshman) students one year or more before high school graduation? [ ] Yes	 [x] No   


## C21-C22: Early Decision and Early Action Plans

### C21. Early Decision

Does your institution offer an early decision plan for first-time, first-year (freshman) applicants for fall enrollment? [ ] Yes  [x] No   

### C22.Early Action

Do you have a nonbinding early action plan whereby students are notified of an admission decision well	in advance of the regular notification date but do not have to commit to attending your college? [ ] Yes  [x] No 


# D. TRANSFER ADMISSION

## D1-D2: Fall Applicants

###  D1. Does your institution enroll transfer students?

- [x] Yes 	
- [ ] No

May transfer students earn advanced standing credit by transferring credits earned from course work completed at other colleges/universities? 
- [x] Yes 	
- [ ] No 


### D2. The number of students who applied, were admitted, and enrolled as degree-seeking transfer students

```{r echo=FALSE}

rm(HSGPA_Mean, HSGPA_Percent)

Applicants <-table(Transfer_App$CDSsex)
Admitted <- table(Transfer_App$CDSsex, Transfer_App$AdmissionStatus)[ ,1]

Transfer_Enr <- subset(ERS_Enr, EnrollmentStatus == 4 & StudentLevelCode < 5)
Enrolled <- table(Transfer_Enr$CDSsex)

Transfer <- cbind(Applicants,Admitted,Enrolled)
rm(Applicants, Admitted, Enrolled)

D2 <- rbind(Transfer, colSums(Transfer))
rm(Transfer, Transfer_App, Transfer_Enr)

rownames(D2) <- c("Men","Women","Total")

kable(D2)
```

## D3-D11: Application for Admission
### D3. terms for which transfers may enroll:

- [x] Fall 	- [ ] Winter - [x] Spring 	- [ ] Summer 

### D4. Must a transfer applicant have a minimum number of credits completed or else must apply as an entering freshman? 

- [x] Yes 	
- [ ] No

the minimum number of credits and the unit of measure?  60 

### D5. Required items for transfer students to apply for admission

Item                  | Required | Recommended | Required Some | Not Required
--------------------- | -------- | ----------- | ------------- | -------------- 
Highschool transcript |          |             |               |             
College transcript(s) |     x    |             |               |            
personal statement    |          |             |               |     x    
Interview             |          |             |               |     x    
Standardized test     |          |             |               |     x    
Statmnt good standing |          |             |               |     x    


### D6. minimum high school grade point average is required of transfer applicants


### D7. minimum college grade point average is required of transfer applicants

2.00

### D8. other application requirements specific to transfer applicants:


### D9. application priority, closing, notification, and candidate reply dates for transfer students.

Semester| Priority Date | Closing Date | Notification Date | Reply Date | Rolling Admission    
--------|---------------|--------------|-------------------|------------|--------------------
Fall    |               |              |                   |            |        x    
Winter  |               |              |                   |            |    
Spring  |               |              |                   |            |    
Summer  |               |              |                   |            |    


### D10. Does an open admission policy apply to transfer students? 

- [ ] Yes 
- [ ] No

### D11. additional requirements for transfer admission


## D12-D17: Transfer Credit Policies

### D12. the lowest grade earned for any course that may be transferred for credit:


### D13. Maximum number of credits or courses that may be transferred from a two-year institution: 

Number: 70 


### D14. Maximum number of credits or courses that may be transferred from a four-year institution: 

Number: 90


### D15. Minimum number of credits that transfers must complete at your institution to earn an associate degree:


### D16. Minimum number of credits that transfers must complete at your institution to earn a bachelor's degree: 


### D17. other transfer credit policies:


## D18-D22: Military Service Transfer Credit Policies


### D18. Does your institution accept the following military/veteran transfer credits:

Credits from:                                   |     Yes    |     No    
------------------------------------------------|------------|------------
American Council on Education (ACE)             |     x      |    
College Level Examination Program (CLEP)        |     x      |     
DANTES Subject Standardized Tests (DSST)        |     x      |     


### D19. Maximum number of credits or courses that may be transferred based on military education evaluated by the American Council on Education (ACE): 


### D20. Maximum number of credits or courses that may be transferred based on Department of Defense supported prior learning assessments (College Level Examination Program (CLEP) or DANTES Subject Standardized Tests (DSST)):


### D21. Are the military/veteran credit transfer policies published on your website?


### D22. Describe other military/veteran transfer credit policies unique to your institution:



# E. ACADEMIC OFFERINGS AND POLICIES


### E1. Special study options:} Identify those programs available at your institution.

programs                                | programs
----------------------------------------|-------------------------------------
[ ] Accelerated program                 | [x] Honors program     
[x] Cooperative education program       | [ ] Independent study     
[ ] Cross-registration                  | [x] Internships     
[x] Distance learning                   | [ ] Liberal arts/career combination     
[x] Double major                        | [ ] Student-designed major     
[ ] Dual enrollment                     | [x] Study abroad     
[ ] Exchange student program (domestic) | [ ] Weekend college     
[ ] External degree program             | 
[x] Other (specify): Service Learning     

### E3. Areas in which all or most students are required to complete some course work prior to graduation:

programs                            | programs
----------------------------------- | -------------------------------------------
[x] Arts/fine arts                  | [x] Humanities     
[ ] Computer literacy               | [x] Mathematics     
[x] English (including composition) | [x] Philosophy     
[ ] Foreign languages               | [x] Sciences (biological or physical)     
[x] History                         | [x] Social science     
[ ] Other(describe):     	
		
# G. ANNUAL EXPENSES		

### G0. The URL of your institution's net price calculator:

http://www.csumentor.edu/finaid/pais/

### G1. Undergraduate full-time tuition, required fees, room and board

Institution Type                                | First-Year | Undergraduates    
------------------------------------------------|------------|----------------
PRIVATE INSTITUTIONS Tuition:                   |            |    
PUBLIC INSTITUTIONS Tuition: In-district:       | \$5,742    | \$5,742    
PUBLIC INSTITUTIONS In-state (out-of-district): | \$5,742    | \$5,742    
PUBLIC INSTITUTIONS Out-of-state:               | \$16,632   | \$16,632    
NONRESIDENT ALIENS Tuition:                     | \$16,632   | \$16,632    
REQUIRED FEES:                                  | \$1,108    | \$1,108    
ROOM AND BOARD: (on-campus)                     | \$15,642   | \$15,642    
ROOM ONLY: (on-campus)                          |            |    
BOARD ONLY: (on-campus meal plan)               |            |    
		
### G2. Number of credits per term a student can take for the stated full-time tuition

- 7 minimum
- 16 maximum    

### G3. Do tuition and fees vary by year of study (e.g., sophomore, junior, senior)?

- [ ] Yes
- [x] No     

### G4. Do tuition and fees vary by undergraduate instructional program?

- [ ] Yes
- [x] No     

### G5. Provide the estimated expenses for a typical full-time undergraduate student:

Expense type		     |Residents|Commuters(living at home)|Commuters(not living at home)    
---------------------|---------------------------------------------------------------------
Books and supplies:  | \$1,948 | \$1,948                 | \$1,948    
Room only:           |         |                         |    
Board only:          |         |                         |     
Room and Board total:|\$15,642 | \$5,248                 | \$14,502    
Transportation:      | \$1,452 | \$1,524                 | \$1,524    
Other expenses:      | \$1,450 | \$1,450                 | \$1,450    
		
### G6. Undergraduate per-credit-hour charges (tuition only):

Institution Type                                | Charge
------------------------------------------------|----------
PRIVATE INSTITUTIONS:                           |    
PUBLIC INSTITUTIONS In-district:                |    
PUBLIC INSTITUTIONS In-state (out-of-district): |    
PUBLIC INSTITUTIONS Out-of-state:               | \$396    
NONRESIDENT ALIENS:                             | \$396    


# H. FINANCIAL AID

## Aid Awarded to Enrolled Undergraduates

### H1. Total dollar amounts awarded to enrolled full-time and less than full-time degree-seeking undergraduates

Indicate the academic year for which data are reported for items H1, H2, H2A, and H6 below:      
- [ ] 2016-2017 estimated or
- [x] 2015-2016 final      
	
	Which needs-analysis methodology does your institution use in awarding institutional aid? (Formerly H3)      
- [x] Federal methodology (FM)      
- [ ] Institutional methodology (IM)      
- [ ] Both FM and IM      	
	
Scholarship type                          |Need-based(\$)|Non-need based(\$)      
------------------------------------------|--------------|-------------------
Scholarships/Grants                       |              |       
Federal                                   |              |      
State                                     |              |      
Institutional                             |              |      
Scholarships/grants from external sources |              |      
Total Scholarships/Grants                 |              |      
------------------------------------------|--------------|-------------------
Self-Help                                 |              |       
Student loans from all sources            |              |      
Federal Work-Study                        |              |       
State and other work study/employment     |              |      
Total Self-help                           |              |      
Parent Loans                              |              |      
Tuition Waivers                           |              |      
Athletic Awards                           |              |      
		
### H2. Number of Enrolled Students Awarded Aid:

Student Type		                                      |FTFTFresh| FT UG  |PT UG       
------------------------------------------------------|---------|------- |-------
a)degree-seeking undergraduate students               |   4401  |  24094 |  10453      
b)applied for need-based financial aid                |   3807  |  18107 |   5796      
c)determined to have financial need                   |   2966  |  16042 |   5108      
d)awarded any financial aid                           |   2936  |  15999 |   5072      
e)awarded any need-based scholarship or grant aid     |   2436  |  12262 |   2408      
f)awarded any need-based self-help aid                |    829  |   5277 |   1118      
g)awarded any non-need-based scholarship or grant aid |    416  |   1361 |    266      
h)need was fully met                                  |   2368  |  12072 |   2318      
i)percentage of need that was met by need-based aid.  |     81% |     75%|     46%      
j)The average financial aid package                   |\$6717   |\$7772  |\$5578      
k)Average need-based scholarship                      |\$3615   |\$3897  |\$2813      
l)Average need-based self-help award                  |\$4036   |\$4503  |\$4184      
m)Average need-based loan                             |\$2640   |\$2208  |\$3104      
		

### H2A. Number of Enrolled Students Awarded Non-need-based Scholarships and Grants:		
		
Student Type		                                             |FTFTFresh| FT UG|PT UG      
-------------------------------------------------------------|---------|------|------
n)institutional non-need-based scholarship or grant aid      |    55   |    97|     4     
o)Average amount of non-need-based scholarship               |\$1287   |\$1683|\$1362      
p)institutional non-need-based athletic scholarship or grant |    62   |    50|     1      
q)Average amount of athletic	scholarships                   |\$2538   |\$1690|\$0      
		
		
### H4. The number of students in the 2016 undergraduate class who started as first-time students and received a bachelor's degree between July 1, 2015 and June 30, 2016. Exclude students who transferred into your institution.


### H5. Number and percent of students in class (defined in H4 above) borrowing from federal, non-federal, and any loan	sources, and the average (or mean) amount borrowed
		
Loan Type		                   |N borrowed|% borrowed |Average borrowed     
-------------------------------|----------|-----------|-----------------
a) Any loan program            |          |    44%    |\$14965      
b) Federal loan programs       |          |    44%    |\$14965      
c) Institutional loan programs |          |           |       
d) State loan programs         |          |           |       
e) Private alternative loans   |          |           |       
		
## Aid to Undergraduate Degree-seeking Nonresident Aliens

### H6. Indicate your institution's policy regarding institutional scholarship and grant aid for undergraduate degree-seeking nonresident aliens:

- [ ] Institutional need-based scholarship or grant aid is available
- [ ] Institutional non-need-based scholarship or grant aid is available
- [ ] Institutional scholarship and grant aid is not available      

- The number of undergraduate degree-seeking nonresident aliens who were awarded need-based or non-need-based aid: 439      
- Average dollar amount of institutional financial aid awarded to undergraduate degree-seeking nonresident aliens: \$3944       
- Total dollar amount of institutional financial aid awarded to undergraduate degree-seeking nonresident aliens: \$4,020,232       

### H7. Financial aid forms nonresident alien first-year financial aid applicants must submit:

- [ ] Institution's own financial aid form
- [ ] CSS/Financial Aid PROFILE
- [ ] International Student's Financial Aid Application
- [ ] International Student's Certification of Finances
- [ ] Other:       
	
## Process for First-Year/Freshman Students	
### H8. Financial aid forms domestic first-year (freshman) financial aid applicants must submit:

- [ ] FAFSA
- [ ] Institution's own financial aid form
- [ ] CSS/Financial Aid PROFILE
- [ ] State aid form
- [ ] Noncustodial PROFILE
- [ ] Business/Farm Supplement
- [ ] Other: 
	
### H9. Filing dates for first-year (freshman) students:

Priority date for filing required financial aid forms:       
Deadline for filing required financial aid forms:       
No deadline for filing required forms (applications processed on a rolling basis):       
	
### H10. notification dates for first-year (freshman) students (answer a or b):      

a.) Students notified on or about (date):       
b.) Students notified on a rolling basis:  - [ ] Yes - [ ] No       
	If yes, starting date:       

### H11. reply dates:      

Students must reply by (date):        or 
within 4  weeks of notification.      
	
	
## Types of Aid Available     
	
### H12. Loans    

- [x]  Direct Subsidized Stafford Loans
- [x]  Direct Unsubsidized Stafford Loans
- [x]  Direct PLUS Loans
- [x]  Federal Perkins Loans
- [ ]  Federal Nursing Loans
- [ ]  State Loans
- [ ]  College/university loans from institutional funds
- [x]  Other (specify): Private alternative loans.      

### H13. Need-Based Scholarships and Grants     

- [x]  Federal Pell
- [x]  SEOG
- [x]  State scholarships/grants
- [x]  Private scholarships
- [ ]  College/university scholarship or grant aid from institutional funds
- [ ]  United Negro College Fund
- [ ]  Federal Nursing Scholarship
- [ ]  Other (specify):   

### H14. Check off criteria used in awarding institutional aid. Check all that apply.

Criteria		             |Non Need Based|Need Based      
-------------------------|--------------|-----------
Academics	               |              |      
Alumni affiliation       |              |      	
Art	                     |              |      
Athletics                |              |      
Job skills	             |              |      
ROTC                     |              |       	
Leadership               |              |      	
Minority status	         |              |      
Music/drama	             |              |      
Religious affiliation    |              |      	
State/district residency |              |      	
		
### H15. If your institution has recently implemented any major financial aid policy, program, or initiative to make your institution more affordable to incoming students such as replacing loans with grants, or waiving costs for families below a certain income level please provide details below:		




		

# J. Didsciplinary areas of DEGREES CONFERRED

## J1. Degrees conferred between July 1, 2020 and June 30, 2021

```{r echo=FALSE}

# Using degree data, select only Bachelors degrees.

Bachelors <- subset(ERSD, DegreeLevelCode < 5)
rm(ERSD)

# Read in COSAR HEGIS CIP Crossover.

CIPlist <- read.fwf(file.path(CIPdir, "commonc-cosarcipcodefxvw-prd-1.csv"),
                  width = c(5,-1,2,-68,1),
                  col.names=c('MajorCode','CIPcode','Obsolete'),
                  colClasses=c('character','character','numeric'))

# Remove Obsolete codes.

CIPxOver <- subset(CIPlist, Obsolete == 0, select = c('MajorCode','CIPcode'))
rm(CIPlist)

# Read in CDS CIP Category names.

CIPcategory <- read.csv(file.path(CIPdir, "CIPCategory.csv"))

# Join the category names to CIP Code.

CIP <- merge(CIPxOver, CIPcategory, by="CIPcode")
rm(CIPxOver)

# Join the CIP code to degree file.

ERSD <- merge(Bachelors, CIP, by="MajorCode")

# Save the frequencies.

J <- data.frame(table(ERSD$Category))
rm(ERSD)

# Calculate the percent and add total.

J$percent <- (J$Freq/(sum(J$Freq)))*100
J$Var1 <- as.character(J$Var1)
J <- rbind(J, c("Total", colSums(J[,2:3, drop=FALSE])))
J$percent <- sprintf("%.1f%%", as.numeric(J$percent))
colnames(J) <- c("Category","Number","Bachelors")

# Reattach CIP Codes.

Jtable <- merge(J, CIPcategory, by="Category", all=TRUE)
rm(J, CIPcategory)

DegreeTable = subset(Jtable, select = -c(Number))
DegreeTable <-DegreeTable[order(DegreeTable$CIPcode, DegreeTable$Category),]
row.names(DegreeTable) <- NULL
rm(Jtable)

kable(DegreeTable)

```
